---
title: "Overview of HEdtree"
author: "P.J Dodd"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Preliminaries

First make sure that the package is loaded:

```{r}
library(HEdtree)
library(data.tree)              #needed?
```

# Making a tree

Then we can use the syntax of `data.tree` to define our decision tree

```{r}
myt <- Node$new('mytree')
ma <- myt$AddChild('Option A')
mb <- myt$AddChild('Option B')
ma1 <- ma$AddChild('Option A1')
ma2 <- ma$AddChild('Option A2')
mb1 <- mb$AddChild('Option B1')
mb2 <- mb$AddChild('Option B2')
```

We can also add incoming probabilities, as well as costs and health-related quality of life parameters for these nodes

```{r}
## p/q/c
myt$p <- 'myp'; myt$qol <- 'myq'; myt$cost <- 'myc'
ma$p <- 'ap'; ma$qol <- 'aq'; ma$cost <- 'ac';
mb$p <- 'bp'; mb$qol <- 'bq'; mb$cost <- 'bc';
ma1$p <- 'a1p'; ma1$qol <- 'a1q'; ma1$cost <- 'a1c';
mb1$p <- 'b1p'; mb1$qol <- 'b1q'; mb1$cost <- 'b1c';
ma2$p <- 'a2p'; ma2$qol <- 'a2q'; ma2$cost <- 'a2c';
mb2$p <- 'b2p'; mb2$qol <- 'b2q'; mb2$cost <- 'b2c';
```

We can check we're on the right lines using ASCI-style visualization:

```{r}
print(myt,'p','cost','qol')
```
If other variables have been attached to nodes, they can be visualized in a similar fashion.


# Calculations

`HEdtree` relies on recursion, closures and string manipulation to perform its calculations. For example, we can generate strings that correspond to the mean cost or HRQoL for entries to the tree.

```{r}
getAQ(myt,'cost')
getAQ(myt,'qol')
```

We can look at an exhaustive list of parameters defined using

```{r}
(varnmz <- showParmz(myt)$vars)
```

The function `showParmz` also collates variables which are defined in terms of other defined parameters (using arithmetic operations e.g.). It is permissable to define a cost using such an operation, e.g. `test$qol <- '(b1q-.1)/2'` 

Functions that evaluate these strings on data.frames are built using

```{r}
H <- makeTfuns(myt) 
```

We can check that these are working by building a dummy PSA data.frame

```{r}
## dummy PSA data frame
psadf <- matrix(runif(10*length(varnmz)),ncol=length(varnmz))
colnames(psadf) <- varnmz
psadf <- as.data.frame(psadf)

## evaluate
H$costfun(psadf)                        #return answers
H$qolfun(psadf)

```

## Use of functions

It is also permitted to call on functions that are defined in higher-level environments, e.g.

```{r}
ma1$p <- 'accpfun(age)'
accpfun <- function(x) exp(x) / (1+exp(x))
psadf$age <- 1:nrow(psadf)
H <- makeTfuns(myt)                   #remake the functions
H$costfun(psadf)                        #test 

```

# Plotting

TODO

# TODO

* individual patient characteristics
* PSA
* subtrees
* looping
* parameter tables


Vignettes are long form documentation commonly included in packages. Because they are part of the distribution of the package, they need to be as compact as possible. The `html_vignette` output type provides a custom style sheet (and tweaks some options) to ensure that the resulting html is as small as possible. The `html_vignette` format:

- Never uses retina figures
- Has a smaller default figure size
- Uses a custom CSS stylesheet instead of the default Twitter Bootstrap style

## Vignette Info

Note the various macros within the `vignette` section of the metadata block above. These are required in order to instruct R how to build the vignette. Note that you should change the `title` field and the `\VignetteIndexEntry` to match the title of your vignette.

## Styles

The `html_vignette` template includes a basic CSS theme. To override this theme you can specify your own CSS in the document metadata as follows:

    output: 
      rmarkdown::html_vignette:
        css: mystyles.css

## Figures

The figure sizes have been customised so that you can easily put two images side-by-side. 

```{r, fig.show='hold'}
plot(1:10)
plot(10:1)
```

You can enable figure captions by `fig_caption: yes` in YAML:

    output:
      rmarkdown::html_vignette:
        fig_caption: yes

Then you can use the chunk option `fig.cap = "Your figure caption."` in **knitr**.

## More Examples

You can write math expressions, e.g. $Y = X\beta + \epsilon$, footnotes^[A footnote here.], and tables, e.g. using `knitr::kable()`.

```{r, echo=FALSE, results='asis'}
knitr::kable(head(mtcars, 10))
```

Also a quote using `>`:

> "He who gives up [code] safety for [code] speed deserves neither."
([via](https://twitter.com/hadleywickham/status/504368538874703872))
