---
title: "Overview of HEdtree"
author: "P.J Dodd"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Preliminaries

First make sure that the package is loaded:

```{r}
library(HEdtree)
library(data.tree)              #needed?
```

# Making a tree

Then we can use the syntax of `data.tree` to define our decision tree by first creating an empty tree object and then adding to it child nodes using the `AddChild()` function.

```{r}
myt <- Node$new('mytree')
ma <- myt$AddChild('Option A')
mb <- myt$AddChild('Option B')
ma1 <- ma$AddChild('Option A1')
ma2 <- ma$AddChild('Option A2')
mb1 <- mb$AddChild('Option B1')
mb2 <- mb$AddChild('Option B2')
```

We can also add incoming probabilities (`p`), as well as costs (`cost`) and health-related quality of life (`qol`) parameters for these nodes

```{r}
## p/q/c
myt$p <- 'myp'; myt$qol <- 'myq'; myt$cost <- 'myc'
ma$p <- 'ap'; ma$qol <- 'aq'; ma$cost <- 'ac'
mb$p <- 'bp'; mb$qol <- 'bq'; mb$cost <- 'bc'
ma1$p <- 'a1p'; ma1$qol <- 'a1q'; ma1$cost <- 'a1c'
mb1$p <- 'b1p'; mb1$qol <- 'b1q'; mb1$cost <- 'b1c'
ma2$p <- 'a2p'; ma2$qol <- 'a2q'; ma2$cost <- 'a2c'
mb2$p <- 'b2p'; mb2$qol <- 'b2q'; mb2$cost <- 'b2c'
```

We can check we're on the right lines using ASCI-style visualization:

```{r}
print(myt,'p','cost','qol')
```

If other variables have been attached to nodes, they can be visualized in a similar fashion.


# Calculations

`HEdtree` relies on recursion, closures and string manipulation to perform its calculations. For example, we can generate strings that correspond to the mean cost or HRQoL for entries to the tree.

```{r}
getAQ(myt,'cost')
getAQ(myt,'qol')
```

We can look at an exhaustive list of parameters defined using

```{r}
(varnmz <- showParmz(myt)$vars)
```

The function `showParmz` also collates variables which are defined in terms of other defined parameters (using arithmetic operations e.g.). It is permissible to define a cost using such an operation, e.g. `r `test$qol <- '(b1q-.1)/2'` 

Functions that evaluate these strings on data.frames are built using

```{r}
H <- makeTfuns(myt) 
```

We can check that these are working by building a dummy PSA data.frame

```{r}
## dummy PSA data frame
psadf <- matrix(runif(10*length(varnmz)), ncol=length(varnmz))
colnames(psadf) <- varnmz
psadf <- as.data.frame(psadf)

## evaluate
H$costfun(psadf)                        #return answers
H$qolfun(psadf)
```

## Use of functions

It is also permitted to call on functions that are defined in higher-level environments, e.g.

```{r}
ma1$p <- 'accpfun(age)'
accpfun <- function(x) exp(x) / (1+exp(x))
psadf$age <- 1:nrow(psadf)          #add an age in 
H <- makeTfuns(myt)                 #remake the functions
H$costfun(psadf)                    #test 
```
Note: this example also illustrates how one can include individual characteristics from a population and use them to influence their outcomes.


# Plotting

Plotting to a web browser can be achieved by

```{r}
plotter(myt)
```

If fewer/more labels are needed, this can be achieved using

```{r}
plotter(myt, varz=c('name'), edgelabel=FALSE)
```

If further formatting or I/O is required, `ToDiagrammeRGraph` exports the graph as a `DiagrammeR` object (see package of the same name). Or just assign the return value to some variable, e.g. `test`, whereupon `test$x$digram` will contain the associated DOT code.

# Reading in parameters

TODO

# Manipulating subtrees


Looping back is not permitted (these are *tree* models after all). However, subtrees can be copied and added, potentiall after modification.

```{r}
mbCopy <- Clone(mb)
mb2$AddChildNode(mbCopy)
print(myt,'p','cost','qol')
```


# TODO

* PSA
* parameter tables



